\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[expansion=false]{microtype}
\usepackage{amsmath,amssymb,booktabs,longtable,array,siunitx,enumitem,hyperref,xurl}
\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue}
\setlist{nosep}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.4em}
\setlength{\LTleft}{0pt}
\setlength{\LTright}{0pt}
\setlength{\tabcolsep}{4pt}
\setlength{\emergencystretch}{2em}
\allowdisplaybreaks
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcommand{\srcref}[1]{\nolinkurl{#1}}
\newcommand{\filepath}[1]{\nolinkurl{#1}}
\newcommand{\code}[1]{\nolinkurl{#1}}
\title{Physics and Mathematics Reference}
\date{Generated 2026-03-19}
\begin{document}
\maketitle

\section{Physical system overview}
The audited execution graph models three coupled analysis layers.

\begin{enumerate}[leftmargin=*]
\item A MATLAB heating model evaluates a discrete operating grid for a resistive nichrome coil inside a heater tube, followed by an aeration-tube network embedded in the vermicomposter bed \srcref{vermicomposter_air_heater_design_model.m:L1172-L1677}; \srcref{vermicomposter_air_heater_design_model.m:L1803-L2248}.
\item A Python summer branch replaces the electrical heater with a spot-cooled plenum and assist blower, then reuses the same reduced-order aeration, evaporation, and two-node bed closures \srcref{heat_transfer_study.py:L2534-L3292}; \srcref{heat_transfer_study.py:L4326-L4438}.
\item A Python annual layer reuses the selected heating and cooling reference points against a representative-year ambient driver \srcref{heat_transfer_study.py:L4547-L5012}; \srcref{heat_transfer_study.py:L5013-L5496}.
\end{enumerate}

The physically important paths are electrical Joule heating of the wire, wire-to-air convection, heater-shell loss to ambient, sensible air-to-bed transfer through the aeration tube wall, direct discharge enthalpy through the perforations, latent heat consumption by evaporation, and bed-to-ambient loss through the vessel wall and top boundary condition.

This document is intentionally more explicit than \filepath{Heat Transfer Study/VermiHeatLoss.tex}. It restates the implemented equation families, identifies where they appear in the repository narrative, and records the exact code locations where each block is implemented or approximated.

\section{Modeling boundaries and scope}
\begin{itemize}[leftmargin=*]
\item MATLAB is the authoritative source of the winter/heating operating-point grid and geometry-driven heater calculations \srcref{vermicomposter_air_heater_design_model.m:L1172-L1677}.
\item Python is authoritative for the final reported heating recommendation, the full summer cooling workflow, the annual layer, and the current text summaries and plots \srcref{heat_transfer_study.py:L2220-L2499}; \srcref{heat_transfer_study.py:L2534-L3292}; \srcref{heat_transfer_study.py:L5013-L7020}.
\item The model is reduced-order. Tubes are one-dimensional marching domains; the bed is a two-node lumped thermal network; plume and habitat calculations are post-processing estimates rather than CFD \srcref{vermicomposter_air_heater_design_model.m:L1803-L2248}; \srcref{heat_transfer_study.py:L1062-L1595}.
\item The summer branch reuses the same bin wall, aeration-tube, evaporation, and two-node closures, but replaces electrical heating with spot-cooled inlet air and an assist-blower static constraint \srcref{heat_transfer_study.py:L2605-L3292}.
\item This reference documents only equations actually found in the repository. Where \filepath{Heat Transfer Study/VermiHeatLoss.tex} contains broader narrative or optional transient discussion not used by the rooted execution graph, that is called out explicitly.
\end{itemize}

\section{Modeling assumptions}
\begin{itemize}[leftmargin=*]
\item Air is treated as an ideal gas with Sutherland-law viscosity and fixed \(c_p\) and \(Pr\) in the property package \srcref{heat_transfer_study.py:L3607-L3675}; \srcref{vermicomposter_air_heater_design_model.m:L2333-L2359}.
\item Heater and aeration tubes are discretized axially and solved by explicit marching rather than simultaneous PDE solution \srcref{vermicomposter_air_heater_design_model.m:L1803-L1999}; \srcref{heat_transfer_study.py:L2707-L2959}.
\item Parallel aeration branches are identical at a given operating point. Total heat, water loss, and power are assembled from per-tube or per-string contributions \srcref{vermicomposter_air_heater_design_model.m:L1497-L1677}.
\item The bed is represented by bottom and top thermal nodes with a single internal conductance between them \srcref{vermicomposter_air_heater_design_model.m:L971-L1075}; \srcref{heat_transfer_study.py:L4326-L4438}.
\item Evaporation is represented as a local latent sink limited by diffusion, carrying capacity, and available sensible heat; there is no resolved liquid transport field in the compost \srcref{vermicomposter_air_heater_design_model.m:L2000-L2122}; \srcref{heat_transfer_study.py:L2567-L2603}.
\item The vented-cover top is a repository-specific mixed boundary condition: insulated cover over most of the top plus free convection over a small vent area \srcref{vermicomposter_air_heater_design_model.m:L971-L1075}; \srcref{Heat Transfer Study/VermiHeatLoss.tex:L785-L847}.
\item Summer spot-cooler behavior is a surrogate model based on configured rated quantities and simple psychrometric limiting, not a first-principles vapor-compression cycle model \srcref{heat_transfer_study.py:L2605-L2705}.
\end{itemize}

\section{Energy balances}
\subsection{Air and psychrometric property package}
Dry-air density is evaluated as
\begin{equation}
\label{eq:air_density}
\rho=\frac{p}{RT}.
\end{equation}
Dynamic viscosity uses the Sutherland form
\begin{equation}
\label{eq:air_viscosity}
\mu=\mu_0\left(\frac{T}{T_0}\right)^{3/2}\frac{T_0+S}{T+S}.
\end{equation}
Thermal conductivity is then reconstructed from
\begin{equation}
\label{eq:air_conductivity}
k=\frac{\mu c_p}{Pr}.
\end{equation}
For the summer psychrometric branch, humidity ratio is computed from vapor partial pressure \(p_v\) as
\begin{equation}
\label{eq:humidity_ratio}
w=0.62198\frac{p_v}{p-p_v},
\end{equation}
and, for a given relative humidity \(\phi\),
\begin{equation}
\label{eq:humidity_ratio_from_rh}
p_v=\phi\,p_{\mathrm{sat}}(T).
\end{equation}

\textbf{Type.} Equation \eqref{eq:air_density} is exact for the ideal-gas approximation adopted here; \eqref{eq:air_viscosity} is a fitted transport-property relation; \eqref{eq:air_conductivity} is a derived closure using fixed \(c_p\) and \(Pr\); \eqref{eq:humidity_ratio}--\eqref{eq:humidity_ratio_from_rh} are exact algebraic psychrometric relations for the implemented moisture basis. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L11-L37}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1089-L1235}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2084-L2245}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L2333-L2359}; \srcref{heat_transfer_study.py:L3607-L3708}. \\
\textbf{Deviation note.} The property package uses constant \(c_p\) and \(Pr\) rather than temperature-dependent tabulated humid-air properties.

\subsection{Tube-flow partitions, perforations, and electrical grouping}
Let the total imposed inlet volumetric flow be \(\dot V_{\mathrm{tot}}\) and the number of parallel aeration branches be \(N_{\mathrm{tubes}}\). The per-tube inlet volumetric flow is
\begin{equation}
\label{eq:volflow_per_tube}
\dot V_{\mathrm{tube,in}}=\frac{\dot V_{\mathrm{tot}}}{N_{\mathrm{tubes}}}.
\end{equation}
Using the inlet density \(\rho_{\mathrm{in}}\), the corresponding per-tube inlet mass flow is
\begin{equation}
\label{eq:massflow_per_tube}
\dot m_{\mathrm{tube,in}}=\rho_{\mathrm{in}}\dot V_{\mathrm{tube,in}}.
\end{equation}
If a fraction \(\phi_{\mathrm{rel}}\) is released through perforations,
\begin{equation}
\label{eq:released_massflow}
\dot m_{\mathrm{rel}}=\phi_{\mathrm{rel}}\dot m_{\mathrm{tube,in}}.
\end{equation}
For \(N_h\) holes of diameter \(d_h\),
\begin{equation}
\label{eq:hole_area}
A_h=\frac{\pi d_h^2}{4},
\qquad
A_{h,\mathrm{tot}}=N_h A_h,
\end{equation}
and the mean release velocity is
\begin{equation}
\label{eq:hole_velocity}
u_h=\frac{\dot V_{\mathrm{rel}}}{N_h A_h}.
\end{equation}

The coil helix geometry is built from the mean coil diameter \(D_{\mathrm{mean}}\), pitch \(p\), and axial coil span \(L_{\mathrm{coil}}\). The per-turn length is
\begin{equation}
\label{eq:helix_turn_length}
L_{\mathrm{turn}}=\sqrt{(\pi D_{\mathrm{mean}})^2+p^2},
\end{equation}
and the wire length is
\begin{equation}
\label{eq:helix_total_length}
L_{\mathrm{wire}}=N_{\mathrm{turns}}L_{\mathrm{turn}},
\qquad
N_{\mathrm{turns}}\approx \frac{L_{\mathrm{coil}}}{p}.
\end{equation}
With wire resistivity \(\rho_{\mathrm{wire}}(T)\) and cross-sectional area \(A_{\mathrm{wire}}\),
\begin{equation}
\label{eq:wire_resistance}
R_{\mathrm{tube}}=\rho_{\mathrm{wire}}(T_{\mathrm{avg}})\frac{L_{\mathrm{wire}}}{A_{\mathrm{wire}}}.
\end{equation}

The current heating implementation supports groups of series-connected tubes in parallel strings. For group \(g\), let \(n_{s,g}\) be the number of tubes in series per string and \(n_{\mathrm{str},g}\) the number of such strings. Then
\begin{equation}
\label{eq:electrical_group_current}
I_{\mathrm{tube},g}=\frac{V}{n_{s,g}R_{\mathrm{tube}}},
\end{equation}
\begin{equation}
\label{eq:electrical_group_power}
P_{\mathrm{tube},g}=I_{\mathrm{tube},g}^2R_{\mathrm{tube}},
\end{equation}
and with \(n_{\mathrm{tubes},g}=n_{s,g}n_{\mathrm{str},g}\),
\begin{equation}
\label{eq:electrical_totals}
I_{\mathrm{tot}}=\sum_g n_{\mathrm{str},g}I_{\mathrm{tube},g},
\qquad
P_{\mathrm{tot}}=\sum_g n_{\mathrm{tubes},g}P_{\mathrm{tube},g}.
\end{equation}
The heater-tube air inlet is the greenhouse ambient state,
\begin{equation}
\label{eq:heater_air_inlet}
T_{\mathrm{air,in}}=T_{\infty},
\end{equation}
and the aeration-tube inlet is the heater-tube outlet,
\begin{equation}
\label{eq:aeration_air_inlet}
T_{\mathrm{air,in,aer}}=T_{\mathrm{air,out,heater}}.
\end{equation}
For any tube of outer diameter \(D_o\), the outside perimeter used to convert an axial length to an external heat-transfer area is
\begin{equation}
\label{eq:outer_perimeter}
P_{\mathrm{out}}=\pi D_o,
\qquad
A_{\mathrm{seg,out}}=P_{\mathrm{out}}\Delta x.
\end{equation}

\textbf{Type.} Equations \eqref{eq:volflow_per_tube}--\eqref{eq:outer_perimeter} are exact algebraic bookkeeping relations on the reduced-order model state. \\
\textbf{Repository-native narrative.} The same physical quantities appear in \filepath{Heat Transfer Study/VermiHeatLoss.tex:L961-L1034}, \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1560-L1738}, and \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2302-L2444}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L763-L875}; \srcref{vermicomposter_air_heater_design_model.m:L1497-L1677}; \srcref{vermicomposter_air_heater_design_model.m:L1684-L1727}. \\
\textbf{Deviation note.} The MATLAB implementation permits mixed string sizes for odd tube counts; the equations above are therefore written in grouped summation form rather than assuming a single equivalent series-parallel pattern.

\subsection{Heater-tube energy balance}
With heater length \(L_{\mathrm{heater}}\) discretized into \(N_{\mathrm{heater}}\) segments,
\begin{equation}
\label{eq:heater_dx}
\Delta x_{\mathrm{heater}}=\frac{L_{\mathrm{heater}}}{N_{\mathrm{heater}}},
\end{equation}
and the local generated electrical heat assigned to segment \(i\) is
\begin{equation}
\label{eq:heater_segment_generation}
q_{\mathrm{gen},i}=P_{\mathrm{branch}}\frac{\Delta x_{\mathrm{heater}}}{L_{\mathrm{heater}}}.
\end{equation}
For heater segment \(i\), the air-energy update is
\begin{equation}
\label{eq:heater_segment_air}
T_{\mathrm{air},i+1}=T_{\mathrm{air},i}+\frac{q_{\mathrm{gen},i}-q_{\mathrm{loss},i}}{\dot m c_p}.
\end{equation}
Wire temperature is then recovered from the local wire-to-air temperature rise
\begin{equation}
\label{eq:wire_delta_t}
\Delta T_{\mathrm{wire},i}=\frac{q_{\mathrm{gen},i}}{h_{\mathrm{wire},i}A'_{\mathrm{wire}}\Delta x},
\qquad
T_{\mathrm{wire},i}=T_{\mathrm{air},i}+\Delta T_{\mathrm{wire},i}.
\end{equation}
The segmentwise heater-shell loss is
\begin{equation}
\label{eq:heater_segment_loss}
q_{\mathrm{loss},i}=U_{\mathrm{heater}}P_{\mathrm{out}}\Delta x_{\mathrm{heater}}\left(T_{\mathrm{air},i}-T_\infty\right).
\end{equation}

\textbf{Type.} Equation \eqref{eq:heater_segment_generation} is exact segment bookkeeping on the reduced model. Equation \eqref{eq:heater_segment_air} is a discretized one-dimensional steady energy balance. Equations \eqref{eq:wire_delta_t} and \eqref{eq:heater_segment_loss} are reduced-order algebraic temperature-drop and shell-loss closures. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L38-L121}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L1803-L1894}; \srcref{heat_transfer_study.py:L3775-L3895}. \\
\textbf{Deviation note.} The wire-to-air temperature difference is represented by a direct convective drop rather than a resolved wire conduction model.

\subsection{Aeration-tube segment balance}
With aeration length \(L_{\mathrm{aer}}\) discretized into \(N_{\mathrm{aer}}\) segments,
\begin{equation}
\label{eq:aeration_dx}
\Delta x_{\mathrm{aer}}=\frac{L_{\mathrm{aer}}}{N_{\mathrm{aer}}},
\end{equation}
and with a uniform segmentwise released mass,
\begin{equation}
\label{eq:released_mass_per_segment}
\dot m_{\mathrm{rel},i}=\frac{\phi_{\mathrm{rel}}\dot m_{\mathrm{tube,in}}}{N_{\mathrm{aer}}}.
\end{equation}
For aeration segment \(i\), the net bed-side energy contribution is
\begin{equation}
\label{eq:aeration_q_net}
q_{\mathrm{bed},i}=q_{\mathrm{wall},i}+q_{\mathrm{jet},i}-q_{\mathrm{evap},i}.
\end{equation}
Summing over the segment set gives
\begin{equation}
\label{eq:aeration_branch_q}
Q_{\mathrm{bed,branch}}=\sum_i q_{\mathrm{bed},i},
\end{equation}
and the total bed heat rate across \(N_{\mathrm{tubes}}\) identical branches is
\begin{equation}
\label{eq:aeration_total_q}
Q_{\mathrm{bed,total}}=N_{\mathrm{tubes}}Q_{\mathrm{bed,branch}}.
\end{equation}
The winter MATLAB march reduces the in-tube mass flow after each segment by
\begin{equation}
\label{eq:aeration_remaining_mdot}
\dot m_{\mathrm{rem},i+1}=\dot m_{\mathrm{rem},i}-\dot m_{\mathrm{rel},i},
\end{equation}
and advances the bulk in-tube air temperature using only the wall-transfer term,
\begin{equation}
\label{eq:aeration_air_march}
T_{\mathrm{air},i+1}=T_{\mathrm{air},i}-\frac{q_{\mathrm{wall},i}}{\dot m_{\mathrm{rem},i+1}c_p}.
\end{equation}
The summer Python branch instead applies the same wall exchange through an exponential segment factor,
\begin{equation}
\label{eq:cooling_wall_factor}
\Theta_i=\exp\!\left[-\frac{U_{\mathrm{aer}}P_{\mathrm{out}}\Delta x_{\mathrm{aer}}}{\dot m_{\mathrm{mean},i}c_p}\right],
\end{equation}
\begin{equation}
\label{eq:cooling_air_after_wall}
T_{\mathrm{air,after\,wall},i}=T_{\mathrm{bed,ref}}+\left(T_{\mathrm{air},i}-T_{\mathrm{bed,ref}}\right)\Theta_i,
\end{equation}
with
\begin{equation}
\label{eq:cooling_q_wall_exact}
q_{\mathrm{wall},i}=\dot m_{\mathrm{mean},i}c_p\left(T_{\mathrm{air},i}-T_{\mathrm{air,after\,wall},i}\right).
\end{equation}

\textbf{Type.} Derived reduced-order balance. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L151-L205}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1586-L1609}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2168-L2238}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L1895-L1999}; \srcref{heat_transfer_study.py:L2821-L2899}. \\
\textbf{Deviation note.} The model treats released-air sensible exchange as a lumped enthalpy source rather than resolving jets in the pore space. The rooted execution graph therefore uses two closely related but not identical march forms: a winter MATLAB explicit wall-only air update and a summer Python exponential wall-factor update.

\subsection{Two-node bed balance}
The bottom and top bed nodes are solved from the steady \(2\times 2\) linear system
\begin{equation}
\label{eq:two_node_matrix}
\begin{bmatrix}
UA_{b,\infty}+UA_{12} & -UA_{12} \\
-UA_{12} & UA_{t,\infty}+UA_{12}
\end{bmatrix}
\begin{bmatrix}
T_b\\
T_t
\end{bmatrix}
=
\begin{bmatrix}
Q_b+UA_{b,\infty}T_\infty\\
Q_t+UA_{t,\infty}T_\infty
\end{bmatrix},
\end{equation}
with
\begin{equation}
\label{eq:two_node_split}
Q_b=f_bQ_{\mathrm{bed,total}},
\qquad
Q_t=(1-f_b)Q_{\mathrm{bed,total}}.
\end{equation}

\textbf{Type.} Derived algebraic reduction of a two-node thermal network. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L365-L424}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2010-L2079}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L2233-L2248}; \srcref{heat_transfer_study.py:L4432-L4438}. \\
\textbf{Deviation note.} The current rooted execution path does not use the optional transient MATLAB PID simulation; the authoritative audit path is the steady algebraic solve above.

\section{Fluid mechanics and pressure-drop model}
The code constructs Reynolds number in the standard form
\begin{equation}
\label{eq:reynolds}
Re=\frac{\rho u D}{\mu}.
\end{equation}
For laminar internal flow, the Darcy friction factor is
\begin{equation}
\label{eq:darcy_laminar}
f=\frac{64}{Re}.
\end{equation}
For turbulent internal flow, the implemented smooth-tube Churchill expression is
\begin{equation}
\label{eq:churchill_friction}
A=\left[2.457\ln\!\left(\frac{1}{(7/Re)^{0.9}}\right)\right]^{16},
\qquad
B=\left(\frac{37530}{Re}\right)^{16},
\end{equation}
\begin{equation}
\label{eq:churchill_friction_full}
f=8\left[\left(\frac{8}{Re}\right)^{12}+\frac{1}{(A+B)^{3/2}}\right]^{1/12}.
\end{equation}
Tube friction losses are accumulated with Darcy--Weisbach:
\begin{equation}
\label{eq:darcy_drop}
\Delta p_f=f\frac{L}{D}\frac{\rho u^2}{2}.
\end{equation}
Abrupt expansion is modeled as
\begin{equation}
\label{eq:sudden_expansion}
\zeta_{\mathrm{exp}}=\left(1-\frac{A_1}{A_2}\right)^2,
\qquad
\Delta p_{\mathrm{exp}}=\zeta_{\mathrm{exp}}\frac{\rho u_1^2}{2}.
\end{equation}
The conical contraction surrogate uses an Idel'chik bellmouth coefficient \(C(\epsilon,\theta)\) and applies
\begin{equation}
\label{eq:idelchik_contraction}
\zeta_{\mathrm{contr}}=C(\epsilon,\theta)\left(\frac{1}{\epsilon}-1\right),
\qquad
\Delta p_{\mathrm{contr}}=\zeta_{\mathrm{contr}}\frac{\rho u_2^2}{2}.
\end{equation}
The distribution-network losses are then assembled as
\begin{equation}
\label{eq:distribution_total_dp}
\Delta p_{\mathrm{dist}}=
\Delta p_{\mathrm{header}}+
\Delta p_{\mathrm{header\rightarrow splitter}}+
\Delta p_{\mathrm{splitter}}+
\Delta p_{\mathrm{connector,fric}}+
\Delta p_{\mathrm{connector\rightarrow branch}}.
\end{equation}
The total branch-plus-distribution drop is
\begin{equation}
\label{eq:system_total_dp}
\Delta p_{\mathrm{tot}}=\Delta p_{\mathrm{branch}}+\Delta p_{\mathrm{dist}}.
\end{equation}

\textbf{Type.} Equations \eqref{eq:reynolds}--\eqref{eq:system_total_dp} are standard fluid-mechanics relations plus a repository-configured minor-loss assembly. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L862-L961}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1899-L2009}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L2123-L2232}; \srcref{vermicomposter_air_heater_design_model.m:L2399-L2407}; \srcref{heat_transfer_study.py:L3755-L3763}; \srcref{heat_transfer_study.py:L3961-L4319}. \\
\textbf{Deviation note.} The distribution hardware beyond the measured geometry is partly represented by editable lumped \(K\)-values, especially the splitter body loss.

\section{Internal convection correlations}
The internal convection coefficient is always formed from
\begin{equation}
\label{eq:internal_h}
h=\frac{Nu\,k}{D}.
\end{equation}
For laminar developing flow, the implemented Hausen-type relation is
\begin{equation}
\label{eq:graetz}
Gz=Re\,Pr\,\frac{D}{L},
\end{equation}
\begin{equation}
\label{eq:internal_laminar_nu}
Nu=3.66+\frac{0.0668\,Gz}{1+0.04\,Gz^{2/3}},
\qquad Re<2300.
\end{equation}
For turbulent flow, the implemented Gnielinski relation is
\begin{equation}
\label{eq:internal_turbulent_nu}
Nu=\max\!\left[
3.66,\,
\frac{(f/8)(Re-1000)Pr}{1+12.7(f/8)^{1/2}(Pr^{2/3}-1)}
\right],
\qquad Re\ge 2300.
\end{equation}

\textbf{Type.} Empirical correlations. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L80-L103}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L862-L961}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L2382-L2398}; \srcref{heat_transfer_study.py:L3764-L3774}. \\
\textbf{Deviation note.} The same branch logic is used for both heater and aeration tubes. No separate roughness correction or transitional blending beyond the branch cutoff is implemented.

\section{External convection correlations}
For crossflow around the heater wire and perforation-scale characteristic diameter, the code uses the Churchill--Bernstein correlation
\begin{equation}
\label{eq:churchill_bernstein}
Nu=
0.3+
\frac{0.62Re^{1/2}Pr^{1/3}}
{\left[1+(0.4/Pr)^{2/3}\right]^{1/4}}
\left[1+\left(\frac{Re}{282000}\right)^{5/8}\right]^{4/5}.
\end{equation}
For external vertical walls, the Rayleigh number is
\begin{equation}
\label{eq:rayleigh}
Ra=\frac{g\beta\Delta T L^3}{\nu\alpha},
\end{equation}
and the Churchill--Chu plate relation is
\begin{equation}
\label{eq:natural_convection_vertical}
Nu=\left[
0.825+
\frac{0.387Ra^{1/6}}
{\left(1+(0.492/Pr)^{9/16}\right)^{8/27}}
\right]^2,
\qquad
h=\frac{Nu k}{L}.
\end{equation}
For horizontal top or vent surfaces, the implemented piecewise free-convection relation is
\begin{equation}
\label{eq:natural_convection_horizontal}
Nu=
\begin{cases}
0.54Ra^{1/4}, & T_s\ge T_\infty,\ Ra\le 10^7,\\
0.15Ra^{1/3}, & T_s\ge T_\infty,\ Ra>10^7,\\
0.27Ra^{1/4}, & T_s<T_\infty.
\end{cases}
\end{equation}

\textbf{Type.} Empirical correlations. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L30-L57}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L203-L244}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L862-L961}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L1111-L1155}; \srcref{vermicomposter_air_heater_design_model.m:L2372-L2381}; \srcref{heat_transfer_study.py:L3710-L3754}. \\
\textbf{Deviation note.} The same Churchill--Bernstein form is reused both for wire cooling and for hole-scale evaporation coupling.

\section{Conduction and thermal resistance networks}
The implemented wall-stack coefficient is
\begin{equation}
\label{eq:bin_wall_u}
U_{\mathrm{wall}}=
\left(
\frac{1}{h_{\mathrm{inside,bin}}}
+\frac{t_{\mathrm{sheet}}}{k_{\mathrm{sheet}}}
+\frac{t_{\mathrm{ins}}}{k_{\mathrm{ins}}}
+\frac{1}{h_{\mathrm{ext}}}
\right)^{-1}.
\end{equation}
The same algebraic structure is used for the heater tube and aeration tube:
\begin{equation}
\label{eq:heater_overall_u}
U_{\mathrm{heater}}=
\left(
\frac{1}{h_{\mathrm{inside,heater}}}
+\frac{D_o-D_i}{2k_{\mathrm{wall}}}
+\frac{t_{\mathrm{ins}}}{k_{\mathrm{ins}}}
+\frac{1}{h_{\mathrm{outside,heater}}}
\right)^{-1},
\end{equation}
\begin{equation}
\label{eq:aeration_overall_u}
U_{\mathrm{aer}}=
\left(
\frac{1}{h_{\mathrm{inside,aer}}}
+\frac{D_o-D_i}{2k_{\mathrm{wall}}}
+\frac{1}{h_{\mathrm{outside,aer}}}
\right)^{-1}.
\end{equation}

\textbf{Type.} Simplified thermal resistance networks. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L103-L200}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L203-L244}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L1102-L1110}; \srcref{vermicomposter_air_heater_design_model.m:L1884-L1894}; \srcref{heat_transfer_study.py:L3737-L3778}. \\
\textbf{Deviation note.} The radial wall terms use thickness-over-conductivity approximations rather than exact cylindrical \(\ln(r_o/r_i)\) resistances. That simplification is in the implementation and is therefore documented here as implemented.

\section{Heater-side overall heat transfer treatment}
The wire-side convective coefficient is
\begin{equation}
\label{eq:wire_h}
h_{\mathrm{wire}}=\frac{Nu_{\mathrm{wire}}k}{D_{\mathrm{wire}}},
\end{equation}
with \(Nu_{\mathrm{wire}}\) given by \eqref{eq:churchill_bernstein}. The heater-shell loss per segment is
\begin{equation}
\label{eq:heater_loss}
q_{\mathrm{loss},i}=U_{\mathrm{heater}}P_o\Delta x\left(T_{\mathrm{air},i}-T_{\infty}\right),
\end{equation}
and the local net air heating is
\begin{equation}
\label{eq:heater_net_air}
q_{\mathrm{to\,air},i}=q_{\mathrm{gen},i}-q_{\mathrm{loss},i}.
\end{equation}

\textbf{Type.} Derived reduced-order heater balance. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L38-L121}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1738-L1898}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L1803-L1894}; \srcref{heat_transfer_study.py:L3775-L3895}. \\
\textbf{Deviation note.} The Python physics re-solver mirrors the MATLAB heater equations for annual and diagnostic use, but the winter operating grid itself still originates from MATLAB.

\section{Aeration-tube heat transfer model}
The wall-mediated sensible contribution is
\begin{equation}
\label{eq:aeration_q_wall}
q_{\mathrm{wall},i}=U_{\mathrm{aer}}P_o\Delta x\left(T_{\mathrm{air},i}-T_{\mathrm{bed,ref}}\right).
\end{equation}
The direct released-air enthalpy term is
\begin{equation}
\label{eq:aeration_q_jet}
q_{\mathrm{jet},i}=\dot m_{\mathrm{rel},i}c_p\left(T_{\mathrm{air},i}-T_{\mathrm{bed,ref}}\right).
\end{equation}
The in-tube air-temperature march is not a single universal formula across the rooted execution graph. The winter MATLAB branch uses the explicit wall-only update in \eqref{eq:aeration_air_march}, whereas the summer Python branch uses the exponential wall-factor form \eqref{eq:cooling_air_after_wall} with the exact wall exchange \eqref{eq:cooling_q_wall_exact}. In both cases, the net bed contribution still follows \eqref{eq:aeration_q_net}, so the sign of \(q_{\mathrm{jet},i}\) and of the total segment contribution naturally reverses between heating and cooling.

\textbf{Type.} Derived reduced-order segment model. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L151-L205}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2168-L2238}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L1895-L1999}; \srcref{heat_transfer_study.py:L2815-L2959}. \\
\textbf{Deviation note.} The air-release fraction is prescribed, not dynamically solved from local perforation pressure and hole discharge coefficients.

\section{Warm-air release / jet-heating treatment}
The released-air term \eqref{eq:aeration_q_jet} is the model's surrogate for direct pore-space heating or cooling by discharged air. It should be interpreted as an enthalpy source or sink attached to the released fraction \(\dot m_{\mathrm{rel},i}\), not as a resolved jet correlation. In the summer branch, the same equation is reused with \(T_{\mathrm{air}}<T_{\mathrm{bed,ref}}\), so \(q_{\mathrm{jet},i}\) becomes negative and contributes to net cooling. The associated radial plume cooling, worm-safe distance, and habitat-exclusion mathematics are documented in full in the later plume/habitat section beginning at \eqref{eq:release_per_length}.

\textbf{Type.} Reduced-order enthalpy model plus repository-specific plume post-processing. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1342-L1585}. \\
\textbf{Implementation.} \srcref{heat_transfer_study.py:L1062-L1306}; \srcref{heat_transfer_study.py:L1686-L1736}. \\
\textbf{Deviation note.} The plume model is not used to modify the main tube-marching solution; it is a separate post-processing estimate.

\section{Moisture, evaporation, and latent-heat treatment}
The current wetted-area basis can be the top plan area, a manual area, or a particle-surface estimate derived from the Table~3 size fractions. For the particle-surface path, the Sauter mean diameter is
\begin{equation}
\label{eq:sauter_diameter}
\frac{1}{d_{32}}=\sum_i\frac{w_i}{d_i},
\end{equation}
so the external surface area per occupied bed volume is
\begin{equation}
\label{eq:external_area_density}
a_{\mathrm{ext}}=\frac{6}{d_{32}}.
\end{equation}
If \(V_{\mathrm{fill}}\) is the filled bed volume and \(f_{\mathrm{access}}\) the accessible wet-surface fraction,
\begin{equation}
\label{eq:wetted_area}
A_{\mathrm{wet}}=f_{\mathrm{access}}a_{\mathrm{ext}}V_{\mathrm{fill}}.
\end{equation}

The saturation property package uses IAPWS ancillary relations. The saturation pressure is represented as
\begin{equation}
\label{eq:psat_iapws}
\ln\!\left(\frac{p_{\mathrm{sat}}}{p_c}\right)=\frac{T_c}{T}\sum_{i=1}^{6}a_i\tau^{b_i},
\qquad
\tau=1-\frac{T}{T_c},
\end{equation}
with analogous ancillary forms for \(\rho_{\ell,\mathrm{sat}}\) and \(\rho_{v,\mathrm{sat}}\). The latent heat is then computed by Clausius--Clapeyron along saturation:
\begin{equation}
\label{eq:latent_heat_cc}
h_{fg}(T)=T\frac{dp_{\mathrm{sat}}}{dT}
\left(
\frac{1}{\rho_{v,\mathrm{sat}}(T)}-\frac{1}{\rho_{\ell,\mathrm{sat}}(T)}
\right).
\end{equation}

The hole-scale convective coefficient is formed from \eqref{eq:churchill_bernstein} and converted to a mass-transfer coefficient by the Lewis relation:
\begin{equation}
\label{eq:h_mass_lewis}
h_m=\frac{h_{\mathrm{evap}}}{\rho c_p Le^{2/3}}.
\end{equation}
The vapor-density driving force is
\begin{equation}
\label{eq:vapor_density_drive}
\Delta\rho_v=\max\!\left[\rho_{v,\mathrm{sat}}(T_s)-\rho_{v,\infty},\,0\right].
\end{equation}
The diffusion-limited, carrying-capacity-limited, and heat-limited evaporation rates are
\begin{equation}
\label{eq:evap_diff}
\dot m_{\mathrm{diff}}=h_mA_{\mathrm{seg}}\Delta\rho_v,
\end{equation}
\begin{equation}
\label{eq:evap_capacity}
\dot m_{\mathrm{cap}}=\dot V_{\mathrm{rel,seg}}\Delta\rho_v,
\end{equation}
\begin{equation}
\label{eq:evap_heat}
\dot m_{\mathrm{heat}}=\frac{q_{\mathrm{avail}}}{h_{fg}},
\qquad
q_{\mathrm{avail}}=\max\!\left[\dot m_{\mathrm{rel,seg}}c_p(T_{\mathrm{air}}-T_{\mathrm{bed}}),\,0\right].
\end{equation}
The implemented segment evaporation rate and latent sink are
\begin{equation}
\label{eq:evap_selected}
\dot m_{\mathrm{evap,seg}}=
\max\!\left[
0,\,
\min\!\left(\dot m_{\mathrm{diff}},\dot m_{\mathrm{cap}},\dot m_{\mathrm{heat}}\right)
\right],
\qquad
q_{\mathrm{evap,seg}}=\dot m_{\mathrm{evap,seg}}h_{fg}.
\end{equation}

\textbf{Type.} Combination of fitted water-property relations, empirical transport relations, and repository-specific area bookkeeping. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1015-L1369}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1586-L1609}; \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2238-L2245}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L891-L937}; \srcref{vermicomposter_air_heater_design_model.m:L2000-L2122}; \srcref{heat_transfer_study.py:L2567-L2603}; \srcref{heat_transfer_study.py:L3607-L3708}. \\
\textbf{Deviation note.} The summer branch computes inlet humidity after spot-cooler condensate removal, so the effective \(\rho_{v,\infty}\) is not always the raw ambient-RH state.

\section{Radial plume, worm-safe stand-off, and habitat-exclusion post-processing}
The released-air post-processing model starts from the released mass flow rate per unit aeration-tube length,
\begin{equation}
\label{eq:release_per_length}
\dot m'_{\mathrm{rel}}=\frac{\phi_{\mathrm{rel}}\dot m_{\mathrm{tube,in}}}{L_{\mathrm{aer}}},
\end{equation}
and the wetted-area density
\begin{equation}
\label{eq:wetted_area_density}
a_{\mathrm{wet}}=\frac{A_{\mathrm{wet}}}{V_{\mathrm{fill}}}.
\end{equation}
The local wet-surface sink temperature is clamped between the configured surface temperature, the bed-reference temperature, and the released-air temperature:
\begin{equation}
\label{eq:plume_sink_temperature}
T_{\mathrm{sink}}=\min\!\left(\max(T_{s,\mathrm{set}},T_{\mathrm{bed,ref}}),T_{\mathrm{air,out}}\right).
\end{equation}
The plume radius is grown as
\begin{equation}
\label{eq:plume_radius}
b(x)=\min\!\left(b_{\max},\,b_0+\alpha x\right),
\end{equation}
so the contacted wetted area per radial length is
\begin{equation}
\label{eq:plume_contact_area}
A'_{\mathrm{wet}}(x)=\pi a_{\mathrm{wet}}b(x)^2.
\end{equation}
The sensible and latent radial sink terms are
\begin{equation}
\label{eq:plume_sensible_sink}
q'_{\mathrm{sens}}(x)=h_{\mathrm{sens}}A'_{\mathrm{wet}}(x)\left(T-T_{\mathrm{sink}}\right),
\end{equation}
\begin{equation}
\label{eq:plume_latent_sink}
q'_{\mathrm{lat}}(x)=\dot m'_{\mathrm{evap}}(x)h_{fg}(T_s),
\end{equation}
and the plume temperature satisfies
\begin{equation}
\label{eq:plume_energy_ode}
\dot m'_{\mathrm{rel}}c_p\frac{dT}{dx}=-
\left[q'_{\mathrm{sens}}(x)+q'_{\mathrm{lat}}(x)\right].
\end{equation}
The local mass-transfer coefficient is formed by the Lewis relation using local air properties,
\begin{equation}
\label{eq:plume_mass_transfer}
h_m(x)=\frac{h_{\mathrm{evap}}}{\rho_{\mathrm{air}}(T)c_{p,\mathrm{air}}\mathrm{Le}^{2/3}},
\end{equation}
which gives the potential diffusion-driven evaporation rate per radial length,
\begin{equation}
\label{eq:plume_evap_potential}
\dot m'_{\mathrm{evap,pot}}=
h_m(x)A'_{\mathrm{wet}}(x)\left[\rho_{v,\mathrm{sat}}(T_s)-\rho_{v,\mathrm{bulk}}(x)\right].
\end{equation}
The code advances the plume humidity explicitly and caps the bulk vapor density at saturation at the updated plume temperature \srcref{heat_transfer_study.py:L1241-L1292}. If the worm-safety threshold is \(T_{\mathrm{safe}}\), the safe radial offset is the first \(x\) such that
\begin{equation}
\label{eq:worm_safe_distance}
T(x)=T_{\mathrm{safe}},
\end{equation}
located in the code by interpolation on the numerically marched profile \srcref{heat_transfer_study.py:L1307-L1332}.

For habitat accounting, the filled-bed height and volume are
\begin{equation}
\label{eq:fill_height}
H_{\mathrm{fill}}=f_{\mathrm{fill}}H,
\qquad
V_{\mathrm{fill}}=LWH_{\mathrm{fill}},
\end{equation}
with filled-bed cross section
\begin{equation}
\label{eq:filled_domain}
\Omega_{\mathrm{fill}}=\left\{(x,y):0\le x\le W,\ 0\le y\le H_{\mathrm{fill}}\right\}.
\end{equation}
If the safe offset is \(x_{\mathrm{safe}}\), the exclusion radius around each tube is
\begin{equation}
\label{eq:exclusion_radius}
R_{\mathrm{exc}}=r_0+x_{\mathrm{safe}}.
\end{equation}
The pipe and exclusion indicator functions are
\begin{equation}
\label{eq:pipe_indicator}
I_{\mathrm{pipe}}(x,y)=
\begin{cases}
1, & \exists i:\ (x-x_i)^2+(y-y_i)^2\le r_0^2,\\
0, & \text{otherwise},
\end{cases}
\end{equation}
\begin{equation}
\label{eq:exclusion_indicator}
I_{\mathrm{exc}}(x,y)=
\begin{cases}
1, & \exists i:\ (x-x_i)^2+(y-y_i)^2\le R_{\mathrm{exc}}^2,\\
0, & \text{otherwise}.
\end{cases}
\end{equation}
The overlap-aware union areas are
\begin{equation}
\label{eq:pipe_union_area}
A_{\mathrm{pipe,union}}=\iint_{\Omega_{\mathrm{fill}}} I_{\mathrm{pipe}}(x,y)\,dA,
\end{equation}
\begin{equation}
\label{eq:exclusion_union_area}
A_{\mathrm{exc,union}}=\iint_{\Omega_{\mathrm{fill}}} I_{\mathrm{exc}}(x,y)\,dA,
\end{equation}
with volumes obtained by extrusion over aeration-tube length,
\begin{equation}
\label{eq:pipe_exclusion_volumes}
V_{\mathrm{pipe}}=A_{\mathrm{pipe,union}}L_{\mathrm{aer}},
\qquad
V_{\mathrm{exc}}=A_{\mathrm{exc,union}}L_{\mathrm{aer}}.
\end{equation}
The shell-only unsafe volume and remaining habitable volume are then
\begin{equation}
\label{eq:unsafe_shell_volume}
V_{\mathrm{unsafe,shell}}=V_{\mathrm{exc}}-V_{\mathrm{pipe}},
\end{equation}
\begin{equation}
\label{eq:habitable_volume}
V_{\mathrm{hab}}=V_{\mathrm{fill}}-V_{\mathrm{unsafe,total}}.
\end{equation}
When the radial plume reaches the left, right, or bottom wall, the code computes an inside-wall temperature from the same wall stack used in the main bin-loss model,
\begin{equation}
\label{eq:wall_return_flux}
q''_{\mathrm{wall}}=
\frac{T_{\mathrm{plume,wall}}-T_\infty}
{\frac{1}{h_{\mathrm{in}}}+\frac{t_{\mathrm{sheet}}}{k_{\mathrm{sheet}}}+\frac{t_{\mathrm{ins}}}{k_{\mathrm{ins}}}+\frac{1}{h_{\mathrm{ext}}}},
\end{equation}
\begin{equation}
\label{eq:wall_inner_temperature}
T_{\mathrm{wall,in}}=T_{\mathrm{plume,wall}}-\frac{q''_{\mathrm{wall}}}{h_{\mathrm{in}}},
\end{equation}
then converts that to an added unsafe band depth with a linear wall-normal interpolation between the bed temperature and the wall-contact temperature \srcref{heat_transfer_study.py:L1531-L1555}. The final reported unsafe volume is the union of plume exclusion cylinders and wall-return bands:
\begin{equation}
\label{eq:unsafe_total_volume}
V_{\mathrm{unsafe,total}}=V_{\mathrm{union}}.
\end{equation}
For overlap reporting, the naive no-overlap sums are
\begin{equation}
\label{eq:naive_overlap_sums}
A_{\mathrm{pipe,sum}}=N_{\mathrm{tubes}}\pi r_0^2,
\qquad
A_{\mathrm{exc,sum}}=N_{\mathrm{tubes}}\pi R_{\mathrm{exc}}^2,
\end{equation}
with overlap corrections
\begin{equation}
\label{eq:overlap_corrections}
A_{\mathrm{pipe,ov}}=A_{\mathrm{pipe,sum}}-A_{\mathrm{pipe,union}},
\qquad
A_{\mathrm{exc,ov}}=A_{\mathrm{exc,sum}}-A_{\mathrm{exc,union}}.
\end{equation}
Numerically, the union areas are evaluated on a midpoint grid:
\begin{equation}
\label{eq:midpoint_union_area}
A\approx \left(\frac{WH_{\mathrm{fill}}}{N_g^2}\right)\sum_{j=1}^{N_g}\sum_{k=1}^{N_g} I(x_j^\ast,y_k^\ast).
\end{equation}

\textbf{Type.} Repository-specific radial thermal/moisture post-processing and overlap-aware geometry bookkeeping. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1277-L1571}. \\
\textbf{Implementation.} \srcref{heat_transfer_study.py:L1119-L1367}; \srcref{heat_transfer_study.py:L1440-L1664}. \\
\textbf{Deviation note.} The plume model is not part of the core heater or cooling-march solve. It is a post-processing estimator used for the radial cooling plot and the habitat-exclusion metrics.

\section{Bin-wall and ambient heat-loss model}
The vessel geometry is
\begin{equation}
\label{eq:bin_areas}
A_{\mathrm{side}}=2HL+2HW,
\qquad
A_{\mathrm{bottom}}=LW,
\qquad
A_{\mathrm{top}}=LW.
\end{equation}
With \(N_{\mathrm{vent}}\) top vent holes of diameter \(d_{\mathrm{vent}}\),
\begin{equation}
\label{eq:vent_area}
A_{\mathrm{vent}}=N_{\mathrm{vent}}\frac{\pi d_{\mathrm{vent}}^2}{4},
\qquad
A_{\mathrm{top,cov}}=A_{\mathrm{top}}-A_{\mathrm{vent}}.
\end{equation}
For open-top convection, the characteristic length is taken as
\begin{equation}
\label{eq:top_characteristic_length}
L_{c,\mathrm{top}}=\frac{A_{\mathrm{top}}}{P_{\mathrm{top}}},
\qquad
P_{\mathrm{top}}=2(L+W),
\end{equation}
so the top natural-convection coefficient is
\begin{equation}
\label{eq:top_convection_h}
h_{\mathrm{top}}=\frac{Nu_{\mathrm{top}}k}{L_{c,\mathrm{top}}},
\end{equation}
with \(Nu_{\mathrm{top}}\) given by \eqref{eq:natural_convection_horizontal}. The total vessel heat-loss forms can then be written explicitly as
\begin{equation}
\label{eq:q_loss_closed}
Q_{\mathrm{loss,closed}}=
U_{\mathrm{wall}}(A_{\mathrm{side}}+A_{\mathrm{bottom}}+A_{\mathrm{top}})(T_{\mathrm{bin}}-T_\infty),
\end{equation}
\begin{equation}
\label{eq:q_loss_open}
Q_{\mathrm{loss,open}}=
U_{\mathrm{wall}}(A_{\mathrm{side}}+A_{\mathrm{bottom}})(T_{\mathrm{bin}}-T_\infty)
+h_{\mathrm{top}}A_{\mathrm{top}}(T_{\mathrm{bin}}-T_\infty).
\end{equation}
For an open top, the top-node and total ambient conductances are
\begin{equation}
\label{eq:topmode_open}
UA_{\mathrm{top,open}}=
U_{\mathrm{wall}}A_{\mathrm{top,wallshare}}
+h_{\mathrm{top}}A_{\mathrm{top}},
\end{equation}
\begin{equation}
\label{eq:ua_total_open}
UA_{\mathrm{tot,open}}=
U_{\mathrm{wall}}(A_{\mathrm{side}}+A_{\mathrm{bottom}})
+h_{\mathrm{top}}A_{\mathrm{top}}.
\end{equation}
For a covered top with a small vent,
\begin{equation}
\label{eq:topmode_vented}
UA_{\mathrm{top,vented}}=
U_{\mathrm{wall}}(A_{\mathrm{top,cov}}+A_{\mathrm{top,wallshare}})
+h_{\mathrm{top,vent}}A_{\mathrm{vent}},
\end{equation}
\begin{equation}
\label{eq:ua_total_vented}
UA_{\mathrm{tot,vented}}=
U_{\mathrm{wall}}(A_{\mathrm{side}}+A_{\mathrm{bottom}}+A_{\mathrm{top,cov}})
+h_{\mathrm{top,vent}}A_{\mathrm{vent}}.
\end{equation}
For a fully covered top,
\begin{equation}
\label{eq:topmode_closed}
UA_{\mathrm{tot,closed}}=
U_{\mathrm{wall}}(A_{\mathrm{side}}+A_{\mathrm{bottom}}+A_{\mathrm{top}}).
\end{equation}
The lumped required heating or cooling load is
\begin{equation}
\label{eq:lumped_required_heat}
Q_{\mathrm{req}}=UA_{\mathrm{tot}}\left(T_{\mathrm{set}}-T_\infty\right).
\end{equation}
The characteristic length and Biot estimate are
\begin{equation}
\label{eq:characteristic_length}
L_c=\frac{V_{\mathrm{fill}}}{A_{\mathrm{side}}+A_{\mathrm{bottom}}+A_{\mathrm{top}}},
\qquad
Bi=\frac{h_{\mathrm{ext}}L_c}{k_{\mathrm{bed}}}.
\end{equation}
The lumped time constant is
\begin{equation}
\label{eq:lumped_tau}
\tau=\frac{C}{UA_{\mathrm{tot}}},
\end{equation}
and the associated cooldown curve is
\begin{equation}
\label{eq:lumped_cooldown}
T(t)=T_\infty+\left(T_0-T_\infty\right)e^{-t/\tau}.
\end{equation}

\textbf{Type.} Derived thermal-network relations plus a repository-specific vented-cover extension. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L603-L847}. \\
\textbf{Implementation.} \srcref{vermicomposter_air_heater_design_model.m:L971-L1171}; \srcref{heat_transfer_study.py:L4326-L4431}. \\
\textbf{Deviation note.} The rooted execution path only plots the lumped cooldown curve when \(Bi<0.1\); the equation exists regardless, but the code treats the plot as conditionally valid rather than unconditional.

\subsection{Summer spot-cooler and assist-blower surrogate}
In the summer branch, the requested sensible cooling duty for target supply temperature \(T_{\mathrm{supply,target}}\) is
\begin{equation}
\label{eq:spot_requested_load}
\dot Q_{\mathrm{req,cool}}=
\max\!\left[\dot m c_p\left(T_{\mathrm{amb}}-T_{\mathrm{supply,target}}\right),\,0\right].
\end{equation}
The actual sensible cooling applied by the spot cooler is capped at the configured or derived sensible capacity:
\begin{equation}
\label{eq:cooling_capacity}
\dot Q_{\mathrm{actual,cool}}=
\min\!\left(\dot Q_{\mathrm{req,cool}},\dot Q_{\mathrm{max,cool}}\right),
\qquad
T_{\mathrm{supply}}=
T_{\mathrm{amb}}-\frac{\dot Q_{\mathrm{actual,cool}}}{\dot m c_p}.
\end{equation}
If a coefficient of performance is supplied,
\begin{equation}
\label{eq:spot_cop_power}
P_{\mathrm{spot}}=\frac{\dot Q_{\mathrm{actual,cool}}}{\mathrm{COP}}.
\end{equation}
The aeration-network inlet is then reset to the actual spot-cooler supply state,
\begin{equation}
\label{eq:summer_air_inlet}
T_{\mathrm{air,in}}=T_{\mathrm{supply,actual}},
\end{equation}
while the electrical heater is disabled:
\begin{equation}
\label{eq:summer_zero_heater}
P_{\mathrm{tot}}=0,
\qquad
I_{\mathrm{tot}}=0.
\end{equation}

The post-cooler condensate calculation uses the dry-air flow rate \(\dot m_{\mathrm{dry}}\) and the saturated outlet humidity ratio \(w_{\mathrm{sat,out}}\):
\begin{equation}
\label{eq:psychrometric_condensate}
\dot m_{\mathrm{cond,psych}}=
\max\!\left[\dot m_{\mathrm{dry}}\left(w_{\mathrm{in}}-w_{\mathrm{sat,out}}\right),\,0\right].
\end{equation}
If a rated condensate capacity \(\dot m_{\mathrm{cond,rated}}\) is provided, the applied condensate rate is
\begin{equation}
\label{eq:condensate_limited}
\dot m_{\mathrm{cond}}=
\min\!\left(\dot m_{\mathrm{cond,psych}},\dot m_{\mathrm{cond,rated}}\right),
\end{equation}
otherwise \(\dot m_{\mathrm{cond}}=\dot m_{\mathrm{cond,psych}}\). The supply humidity ratio is then
\begin{equation}
\label{eq:outlet_humidity_ratio}
w_{\mathrm{out}}=
\max\!\left(w_{\mathrm{in}}-\frac{\dot m_{\mathrm{cond}}}{\dot m_{\mathrm{dry}}},\,0\right).
\end{equation}

The assist blower is represented by a linear static curve. With shutoff pressure \(\Delta p_{\mathrm{shutoff}}\), rated-duty point \((\dot V_{\mathrm{rated}},\Delta p_{\mathrm{rated}})\), and implied free-air flow \(\dot V_{\mathrm{free}}\),
\begin{equation}
\label{eq:assist_blower_free_air}
\dot V_{\mathrm{free}}=
\dot V_{\mathrm{rated}}
\frac{\Delta p_{\mathrm{shutoff}}}{\Delta p_{\mathrm{shutoff}}-\Delta p_{\mathrm{rated}}},
\end{equation}
\begin{equation}
\label{eq:assist_blower_linear_curve}
\Delta p_{\mathrm{avail}}(\dot V)=
\max\!\left[
\Delta p_{\mathrm{shutoff}}
\left(1-\frac{\dot V}{\dot V_{\mathrm{free}}}\right),\,0
\right].
\end{equation}
The total summer water-loss rate is
\begin{equation}
\label{eq:summer_water_loss_day}
\dot m_{\mathrm{water,day}}=
86400\,N_{\mathrm{tubes}}\sum_i \dot m_{\mathrm{evap},i},
\end{equation}
the reported net cooling is
\begin{equation}
\label{eq:summer_cooling_capacity}
Q_{\mathrm{cool}}=\max\!\left(-Q_{\mathrm{bed,total}},\,0\right),
\end{equation}
and the corresponding lumped cooling requirement at the cooling setpoint is
\begin{equation}
\label{eq:summer_required_cooling}
Q_{\mathrm{req,cool}}=UA_{\mathrm{tot}}\left(T_{\infty,\mathrm{summer}}-T_{\mathrm{set,cool}}\right).
\end{equation}

\textbf{Type.} Repository-specific surrogate model. Equations \eqref{eq:spot_requested_load}--\eqref{eq:spot_cop_power} are simplified sensible-capacity relations; \eqref{eq:psychrometric_condensate}--\eqref{eq:outlet_humidity_ratio} are psychrometric mass-balance relations; \eqref{eq:assist_blower_free_air}--\eqref{eq:assist_blower_linear_curve} are linearized blower approximations. \\
\textbf{Repository-native narrative.} \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2084-L2284}. \\
\textbf{Implementation.} \srcref{heat_transfer_study.py:L2534-L2565}; \srcref{heat_transfer_study.py:L2605-L2705}. \\
\textbf{Deviation note.} The blower curve is a two-point linear approximation, not a full manufacturer fan map, and the spot-cooler model does not include a detailed refrigeration-cycle calculation.

\section{Constraints and feasibility conditions}
The current heating hard-constraint residual vector is
\begin{equation}
\label{eq:heating_constraint_residuals}
\mathbf{g}_{\mathrm{heat}}=
\begin{bmatrix}
\max(0,I-I_{\max})\\
\max(0,T_{\mathrm{wire}}-T_{\mathrm{wire,max}})\\
\max(0,T_{\mathrm{air,out}}-T_{\mathrm{air,out,max}})\\
\max(0,\Delta p-\Delta p_{\max})\\
\max(0,\dot m_{\mathrm{water}}-\dot m_{\mathrm{water,max}})\\
\max(0,u_h-u_{h,\max})\\
\max(0,T_{b,\min}-T_b)\\
\max(0,T_{t,\min}-T_t)\\
\max(0,T_b-T_{b,\max})\\
\max(0,T_t-T_{t,\max})\\
\max(0,|T_b-T_t|-\Delta T_{\max})\\
\max(0,\mathrm{dutyNeeded}-1)\\
\max(0,-Q_{\mathrm{bed,total}})
\end{bmatrix}.
\end{equation}
The current cooling hard-constraint residual vector is
\begin{equation}
\label{eq:cooling_constraint_residuals}
\mathbf{g}_{\mathrm{cool}}=
\begin{bmatrix}
\max(0,\Delta p-\Delta p_{\mathrm{avail}})\\
\max(0,\Delta p-\Delta p_{\max})\\
\max(0,\dot m_{\mathrm{water}}-\dot m_{\mathrm{water,max}})\\
\max(0,u_h-u_{h,\max})\\
\max(0,T_{b,\min}-T_b)\\
\max(0,T_{t,\min}-T_t)\\
\max(0,T_b-T_{b,\max})\\
\max(0,T_t-T_{t,\max})\\
\max(0,|T_b-T_t|-\Delta T_{\max})\\
\max(0,-Q_{\mathrm{cool}})
\end{bmatrix},
\end{equation}
where \(Q_{\mathrm{cool}}=\max(-Q_{\mathrm{bed,total}},0)\).

\textbf{Type.} Repository-defined feasibility residuals. \\
\textbf{Repository-native narrative.} Heating constraints appear in \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1614-L1738}. The historical summer lexicographic block at \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2251-L2279} is intentionally \emph{not} treated as governing because the live Python code now uses constrained NSGA-II for cooling. \\
\textbf{Implementation.} \srcref{heat_transfer_study.py:L2220-L2451}; \srcref{heat_transfer_study.py:L2963-L3147}. \\
\textbf{Deviation note.} The cooling NSGA-II constraints require positive cooling capacity, but they do \emph{not} currently enforce \(Q_{\mathrm{cool}}\ge Q_{\mathrm{req,cool}}\). Cooling shortfall is computed for diagnostics, not as a hard feasibility constraint.

\section{Objective functions and optimization quantities}
Heating now uses constrained NSGA-II on the discrete candidate grid with objective vector
\begin{equation}
\label{eq:heating_objective_vector}
\mathbf{f}_{\mathrm{heat}}=
\begin{bmatrix}
P_{\mathrm{tot}}\\
-\min(T_b,T_t)\\
|T_b-T_t|
\end{bmatrix}.
\end{equation}
The reported heating design is then selected from the feasible Pareto set by the report-priority tuple
\begin{equation}
\label{eq:reported_design_selection}
\pi_{\mathrm{heat,report}}=
\begin{bmatrix}
-\min(T_b,T_t)\\
P_{\mathrm{tot}}\\
|T_b-T_t|\\
-\tfrac{1}{2}(T_b+T_t)\\
\max(0,\mathrm{dutyNeeded}-1)
\end{bmatrix}.
\end{equation}

Cooling now uses constrained NSGA-II as well. The default cooling objective vector is
\begin{equation}
\label{eq:cooling_objective_vector}
\mathbf{f}_{\mathrm{cool}}=
\begin{bmatrix}
\max(T_b,T_t)\\
\tfrac{1}{2}(T_b+T_t)\\
P_{\mathrm{spot}}+P_{\mathrm{blower}}
\end{bmatrix},
\end{equation}
and the reported cooling design is selected from the feasible cooling Pareto set by
\begin{equation}
\label{eq:cooling_report_selection}
\pi_{\mathrm{cool,report}}=
\begin{bmatrix}
\max(T_b,T_t)\\
\tfrac{1}{2}(T_b+T_t)\\
P_{\mathrm{spot}}+P_{\mathrm{blower}}\\
\dot V_{\mathrm{tot}}\\
|T_b-T_t|\\
-Q_{\mathrm{cool}}
\end{bmatrix}.
\end{equation}
If no feasible cooling Pareto point exists, the code falls back to a best-available diagnostic ranking using the configured ordered tuple
\begin{equation}
\label{eq:cooling_fallback_selection}
\pi_{\mathrm{cool,fallback}}=
\left[
g_{\sigma(1)},g_{\sigma(2)},\dots,g_{\sigma(m)}
\right],
\end{equation}
where \(\sigma\) is the configured \code{priority\_order} permutation over the implemented cooling metrics and residuals \srcref{heat_transfer_study.py:L3133-L3149}; \srcref{heat_transfer_study.py:L3173-L3177}; \srcref{heat_transfer_study.py:L3244-L3246}. This is a live fallback/reporting rule, not the stale historical summer primary-selection rule described in \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2251-L2279}.

\textbf{Type.} Repository-defined optimization quantities. \\
\textbf{Repository-native narrative.} Heating NSGA-II is described in \filepath{Heat Transfer Study/VermiHeatLoss.tex:L1614-L1738}. The stale summer lexicographic primary-selection narrative in \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2251-L2279} is intentionally excluded from the governing-equation set here because it is not implemented. \\
\textbf{Implementation.} \srcref{heat_transfer_study.py:L51-L121}; \srcref{heat_transfer_study.py:L2181-L2499}; \srcref{heat_transfer_study.py:L3148-L3208}; \srcref{heat_transfer_study.py:L6220-L6490}. \\
\textbf{Deviation note.} NSGA-II is used on an already evaluated discrete set. It is not replacing the physics solve with a continuous surrogate optimizer.

\section{Numerical approximations, interpolation, and discretization}
\begin{itemize}[leftmargin=*]
\item The heating design space is generated by exhaustive discrete sweep in MATLAB, not by direct continuous optimization \srcref{vermicomposter_air_heater_design_model.m:L1172-L1476}.
\item Electrical resistance depends on average wire temperature, so each operating point uses a short fixed-point iteration on \(T_{\mathrm{wire,avg}}\) \srcref{vermicomposter_air_heater_design_model.m:L1497-L1536}.
\item Heater and aeration transport are solved by explicit axial marching over configured segment counts \srcref{vermicomposter_air_heater_design_model.m:L1803-L1999}; \srcref{heat_transfer_study.py:L2707-L2959}; \srcref{heat_transfer_study.py:L3775-L3959}.
\item The summer branch solves for a bed-reference temperature by bracketing and bisection over the scalar residual relating branch cooling to the lumped required cooling target \srcref{heat_transfer_study.py:L2774-L2899}.
\item Display interpolation in the plotting layer is presentation-only and does not alter the underlying evaluated candidate set \srcref{heat_transfer_study.py:L555-L562}.
\item The annual climate layer uses periodic Gaussian-kernel regression over day-of-year and then overlays freeze-day and heat-wave events \srcref{heat_transfer_study.py:L4633-L4858}.
\item Radial plume and habitat calculations use midpoint-grid union-area estimates rather than closed-form geometric intersection formulas \srcref{heat_transfer_study.py:L1383-L1595}.
\item NSGA-II is applied to integer-valued indices or discrete point indices through \texttt{pymoo}; the optimizer explores the discrete repository candidate set rather than an interpolated continuous space \srcref{heat_transfer_study.py:L2410-L2499}; \srcref{heat_transfer_study.py:L3123-L3208}.
\end{itemize}

\subsection*{Detailed marching and fixed-point closures}
The current heating solver uses the temperature-dependent wire resistivity
\begin{equation}
\label{eq:wire_resistivity_temperature}
\rho_{\mathrm{wire}}(T)=\rho_{\mathrm{wire},20}\left[1+\alpha_{\mathrm{wire}}(T-20^\circ\mathrm{C})\right],
\end{equation}
which yields a tube or branch resistance of the form already introduced in \eqref{eq:wire_resistance}. The fixed-point iteration updates the average wire temperature by under-relaxation:
\begin{equation}
\label{eq:wire_underrelaxation}
\overline{T}_{\mathrm{wire}}^{(k+1)}
=0.65\,\overline{T}_{\mathrm{wire}}^{(k)}+0.35\,\overline{T}_{\mathrm{wire,new}}^{(k)},
\end{equation}
and stops when
\begin{equation}
\label{eq:wire_convergence}
\left|\overline{T}_{\mathrm{wire,new}}^{(k)}-\overline{T}_{\mathrm{wire}}^{(k)}\right|<0.05^\circ\mathrm{C},
\end{equation}
or when the configured iteration cap is reached \srcref{vermicomposter_air_heater_design_model.m:L1497-L1536}. The winter aeration pressure-drop march uses the segment contribution
\begin{equation}
\label{eq:segment_pressure_drop}
\Delta p_{\mathrm{seg}}=
\left(f\frac{\Delta x}{D}+\frac{K_{\mathrm{minor}}}{N}\right)\frac{\rho u^2}{2},
\end{equation}
with the total candidate pressure drop assembled from the marched branch losses plus the distribution-network losses in \eqref{eq:distribution_total_dp} \srcref{vermicomposter_air_heater_design_model.m:L1953-L1956}; \srcref{vermicomposter_air_heater_design_model.m:L2123-L2232}.

The summer branch closes the bed-reference temperature by solving the scalar consistency condition
\begin{equation}
\label{eq:summer_bed_reference_residual}
R(T_{\mathrm{bed,ref}})=\frac{T_b(T_{\mathrm{bed,ref}})+T_t(T_{\mathrm{bed,ref}})}{2}-T_{\mathrm{bed,ref}}=0,
\end{equation}
where \(T_b\) and \(T_t\) are obtained from the two-node bed solve using the branch heat rate predicted at that trial reference temperature \srcref{heat_transfer_study.py:L2910-L2917}. If the lower and upper trial bounds bracket a sign change, the code applies bisection:
\begin{equation}
\label{eq:summer_bisection_midpoint}
T_{\mathrm{mid}}=\frac{T_{\mathrm{lo}}+T_{\mathrm{hi}}}{2},
\end{equation}
updating the bracket by the sign of \(R(T_{\mathrm{mid}})\) until either
\begin{equation}
\label{eq:summer_bisection_stop}
\left|R(T_{\mathrm{mid}})\right|<0.02
\qquad \text{or} \qquad
\left|T_{\mathrm{hi}}-T_{\mathrm{lo}}\right|<0.02,
\end{equation}
otherwise the closer endpoint is retained \srcref{heat_transfer_study.py:L2918-L2947}. These are repository numerical constructs rather than textbook closed forms.

\section{Annual climate and year-round operating layer}
\subsection{Climate anchors and regression baseline}
The annual layer stores for each month \(m\) a Connecticut monthly mean temperature \(\bar T_m\) and monthly standard deviation \(\sigma_m\) derived from the NOAA series described in \filepath{Heat Transfer Study/VermiHeatLoss.tex:L2576-L2616}. The representative-year ambient baseline uses circular Gaussian-kernel regression:
\begin{equation}
\label{eq:kernel_baseline}
\hat y(d)=\frac{\sum_i w_i(d)y_i}{\sum_i w_i(d)},
\qquad
w_i(d)=\exp\!\left[-\frac{1}{2}\left(\frac{\delta_i(d)}{h}\right)^2\right],
\end{equation}
with wraparound day distance
\begin{equation}
\label{eq:circular_day_distance}
\delta_i(d)=\min\!\left(|d-d_i|,\ P-|d-d_i|\right).
\end{equation}
This same periodic Gaussian regression is applied both to the monthly mean anchors and to the monthly standard-deviation anchors \srcref{heat_transfer_study.py:L4794-L4808}; \srcref{heat_transfer_study.py:L4866-L4879}.

\subsection{Freeze-day allocation and anomaly override}
Expected monthly freeze-day counts are converted to representative integer counts by largest-remainder allocation:
\begin{equation}
\label{eq:largest_remainder}
n_{f,m}=\mathrm{LR}(F_m).
\end{equation}
Within each month, day-of-month weights are centered on the observed mean freeze day \(\mu_{f,m}\):
\begin{equation}
\label{eq:freeze_weight}
w_{f,m}(i)=\exp\!\left[-\frac{(i-\mu_{f,m})^2}{2s_{f,m}^2}\right],
\end{equation}
and sequential selection applies a repulsion modifier
\begin{equation}
\label{eq:freeze_repulsion}
\tilde w_{f,m}(i)=
w_{f,m}(i)\prod_{j\in\mathcal{S}_{f,m}}
\left[1-0.65\exp\!\left(-\frac{(i-j)^2}{2r_{f,m}^2}\right)\right].
\end{equation}
On the selected freeze days, the ambient profile is overridden by
\begin{equation}
\label{eq:freeze_override}
T_m(d)=\min\!\left(T_{\mathrm{base},m}(d),\,\bar T_m-\Delta T_{f,m}\right).
\end{equation}
These relations correspond to \srcref{heat_transfer_study.py:L4683-L4732}; \srcref{heat_transfer_study.py:L4893-L4906}.

\subsection{Heat-wave allocation and anomaly override}
Monthly spell starts and total heat-wave days are likewise allocated by largest-remainder rounding:
\begin{equation}
\label{eq:heatwave_alloc}
n_{s,m}=\mathrm{LR}(S_m),
\qquad
n_{h,m}=\mathrm{LR}(H_m).
\end{equation}
The start-day weighting is
\begin{equation}
\label{eq:heatwave_weight}
w_{h,m}(i)=\exp\!\left[-\frac{(i-\mu_{h,m})^2}{2s_{h,m}^2}\right],
\end{equation}
with repulsion-modified score
\begin{equation}
\label{eq:heatwave_repulsion}
\tilde w_{h,m}(i)=
w_{h,m}(i)\prod_{j\in\mathcal{S}_{h,m}}
\left[1-0.65\exp\!\left(-\frac{(i-j)^2}{2r_{h,m}^2}\right)\right].
\end{equation}
Each selected heat-wave spell is forced to last at least three days in the code, and the ambient override is
\begin{equation}
\label{eq:heatwave_override}
T_m(d)=\max\!\left(T_m(d),\,\bar T_m+\Delta T_{h,m}\right).
\end{equation}
These relations correspond to \srcref{heat_transfer_study.py:L4735-L4791}; \srcref{heat_transfer_study.py:L4907-L4911}.

\subsection{Daily heating and cooling loads}
For each representative day \(d\), the required heating and cooling loads are
\begin{equation}
\label{eq:daily_required_heating}
Q_{\mathrm{req},h}(d)=UA_{\mathrm{bin}}\max\!\left(T_{h,\mathrm{set}}-T_\infty(d),\,0\right),
\end{equation}
\begin{equation}
\label{eq:daily_required_cooling}
Q_{\mathrm{req},c}(d)=UA_{\mathrm{bin}}\max\!\left(T_\infty(d)-T_{c,\mathrm{set}},\,0\right),
\end{equation}
which are implemented by rebuilding the bin model at the day's ambient and calling the same lumped-loss routine used elsewhere \srcref{heat_transfer_study.py:L4977-L4983}.

\subsection{Reference-point capacity re-solve and duty application}
The annual layer re-solves the selected heating and cooling operating points at each day’s ambient temperature to obtain
\begin{equation}
\label{eq:annual_heating_capacity}
Q_{h,\mathrm{cap}}(d)=Q_{\mathrm{to\,bed},h}\!\left(T_\infty(d)\right),
\qquad
P_{h,\mathrm{elec}}(d)=P_h\!\left(T_\infty(d)\right),
\end{equation}
\begin{equation}
\label{eq:annual_cooling_capacity}
Q_{c,\mathrm{cap}}(d)=-Q_{\mathrm{to\,bed},c}\!\left(T_\infty(d)\right),
\qquad
P_{c,\mathrm{elec}}(d)=P_{\mathrm{spot\,cooler}}(d)+P_{\mathrm{assist\,blower}}(d).
\end{equation}
The required duty fractions are
\begin{equation}
\label{eq:daily_duty_required}
D_h^\ast(d)=\frac{Q_{\mathrm{req},h}(d)}{Q_{h,\mathrm{cap}}(d)},
\qquad
D_c^\ast(d)=\frac{Q_{\mathrm{req},c}(d)}{Q_{c,\mathrm{cap}}(d)},
\end{equation}
and the applied duties are clipped at unity:
\begin{equation}
\label{eq:daily_duty}
f_{\mathrm{day}}=
\begin{cases}
\min\!\left(\max\!\left(\dfrac{\dot Q_{\mathrm{load}}}{\dot Q_{\mathrm{nom}}},0\right),1\right), & \dot Q_{\mathrm{nom}}>0,\\
0, & \dot Q_{\mathrm{nom}}=0\ \text{and}\ \dot Q_{\mathrm{load}}\le 0,\\
1, & \dot Q_{\mathrm{nom}}=0\ \text{and}\ \dot Q_{\mathrm{load}}>0,
\end{cases}
\end{equation}
which, in mode-specific notation, are exactly
\begin{equation}
\label{eq:daily_duty_clipped}
D_h(d)=\min(\max(D_h^\ast(d),0),1),
\qquad
D_c(d)=\min(\max(D_c^\ast(d),0),1).
\end{equation}
The applied bed load is therefore
\begin{equation}
\label{eq:daily_applied_bed_load}
Q_{\mathrm{bed}}(d)=
\begin{cases}
D_h(d)\,Q_{h,\mathrm{cap}}(d), & \text{heating day},\\
-D_c(d)\,Q_{c,\mathrm{cap}}(d), & \text{cooling day},\\
0, & \text{idle day}.
\end{cases}
\end{equation}

\subsection{Daily bed temperatures, electricity, cost, and water}
The daily bed temperatures are obtained from the same two-node steady system:
\begin{equation}
\label{eq:daily_two_node_matrix}
\begin{bmatrix}
UA_{b,\infty}+UA_{bt} & -UA_{bt}\\
-UA_{bt} & UA_{t,\infty}+UA_{bt}
\end{bmatrix}
\begin{bmatrix}
T_b(d)\\
T_t(d)
\end{bmatrix}
=
\begin{bmatrix}
f_bQ_{\mathrm{bed}}(d)+UA_{b,\infty}T_\infty(d)\\
(1-f_b)Q_{\mathrm{bed}}(d)+UA_{t,\infty}T_\infty(d)
\end{bmatrix}.
\end{equation}
Daily heating and cooling electricity are
\begin{equation}
\label{eq:daily_energy_modes}
E_h(d)=\frac{P_{h,\mathrm{elec}}(d)\,D_h(d)\,24}{1000},
\qquad
E_c(d)=\frac{P_{c,\mathrm{elec}}(d)\,D_c(d)\,24}{1000},
\end{equation}
so the total day electrical energy is
\begin{equation}
\label{eq:daily_energy}
E_{\mathrm{day}}=E_h(d)+E_c(d).
\end{equation}
With electricity price \(c_{\mathrm{elec}}\), the daily cost is
\begin{equation}
\label{eq:daily_cost}
C(d)=c_{\mathrm{elec}}E_{\mathrm{day}},
\end{equation}
and the daily water-replacement estimate is
\begin{equation}
\label{eq:daily_water}
m_{\mathrm{water}}(d)=\dot m_{\mathrm{water},24h}\,D(d),
\end{equation}
where \(D(d)\) is the active heating or cooling duty fraction for that day. These are repository numerical constructs driven by the same daily re-solved heating/cooling reference points \srcref{heat_transfer_study.py:L4966-L5072}.

\section{Implementation notes and deviations from the written model}
\begin{itemize}[leftmargin=*]
\item The authoritative reported heating recommendation is now Python-side, even though MATLAB still generates the heating candidate physics. The rooted execution path therefore treats Python as the final recommendation authority \srcref{heat_transfer_study.py:L6908-L6947}; \srcref{vermicomposter_air_heater_design_model.m:L536-L608}.
\item Maximum bottom and top bed temperatures are enforced in the Python-side heating and cooling feasibility logic; the MATLAB helper \code{explicitlySafe(...)} still omits those maxima \srcref{vermicomposter_air_heater_design_model.m:L2647-L2659}; \srcref{heat_transfer_study.py:L2220-L2368}; \srcref{heat_transfer_study.py:L2963-L3112}.
\item Cooling selection is no longer lexicographic. The current code uses constrained NSGA-II plus a report-priority ordering inside the feasible cooling Pareto set \srcref{heat_transfer_study.py:L3148-L3208}; \srcref{heat_transfer_study.py:L6220-L6490}.
\item Cooling shortfall is computed and printed, but not enforced as a hard cooling feasibility condition in the current NSGA-II implementation \srcref{heat_transfer_study.py:L2915-L2959}; \srcref{heat_transfer_study.py:L2963-L3112}.
\item The post-cooler relative humidity used by the summer evaporation model is recomputed after psychrometric condensate removal, limited by any configured rated condensate capacity \srcref{heat_transfer_study.py:L2662-L2673}.
\item The rooted refresh/export path disables MATLAB plotting and transients in the generated export helper, so the rooted execution graph exercises steady heating physics rather than the optional MATLAB transient branch \srcref{analyze_vermicomposter_heater.py:L294-L304}.
\end{itemize}

\section{Symbol glossary with units}
\begingroup
\small
\begin{longtable}{@{}L{0.16\textwidth}L{0.60\textwidth}L{0.14\textwidth}@{}}
\toprule
Symbol & Meaning & Units \\
\midrule
\endfirsthead
\toprule
Symbol & Meaning & Units \\
\midrule
\endhead
\(A_h\) & area of one perforation hole & \si{\meter\squared} \\
\(A_{\mathrm{wet}}\) & effective wetted area assigned to evaporation & \si{\meter\squared} \\
\(Bi\) & lumped-capacitance Biot number & 1 \\
\(c_p\) & specific heat of air & \si{\joule\per\kilogram\per\kelvin} \\
\(d_{32}\) & Sauter mean diameter used in particle-area estimate & \si{\meter} \\
\(D_i,D_o\) & inner and outer tube diameters & \si{\meter} \\
\(f\) & Darcy friction factor & 1 \\
\(h\) & convective heat-transfer coefficient & \si{\watt\per\meter\squared\per\kelvin} \\
\(h_m\) & mass-transfer coefficient & \si{\meter\per\second} \\
\(h_{fg}\) & latent heat of vaporization & \si{\joule\per\kilogram} \\
\(I_{\mathrm{tot}}\) & total electrical current draw & \si{\ampere} \\
\(Nu,Re,Pr,Ra\) & standard dimensionless groups & 1 \\
\(P_{\mathrm{tot}}\) & total electrical power & \si{\watt} \\
\(Q_{\mathrm{bed,total}}\) & net total bed heat rate; negative in cooling mode & \si{\watt} \\
\(Q_{\mathrm{cool}}\) & positive cooling capacity, \(\max(-Q_{\mathrm{bed,total}},0)\) & \si{\watt} \\
\(Q_{\mathrm{req}}\) & lumped required heating or cooling load & \si{\watt} \\
\(T_b,T_t\) & bottom and top bed-node temperatures & \si{\celsius} \\
\(T_\infty\) & ambient air temperature & \si{\celsius} \\
\(UA\) & lumped conductance & \si{\watt\per\kelvin} \\
\(u_h\) & mean hole exit velocity & \si{\meter\per\second} \\
\(\dot m_{\mathrm{evap}}\) & evaporation mass flow rate & \si{\kilogram\per\second} \\
\(\dot V_{\mathrm{tot}}\) & total imposed volumetric flow rate & \si{\meter\cubed\per\second} \\
\bottomrule
\end{longtable}
\endgroup

\section{Bibliography / source notes}
\begin{itemize}[leftmargin=*]
\item [R1] Churchill and Bernstein cylinder-in-crossflow citation string is stored in \srcref{vermicomposter_air_heater_design_model.m:L4-L8}. The repository preserves the authors, title, journal, year, and page range 300--306; no in-repository PDF was found for deeper page verification.
\item [R2] Gnielinski's internal-flow correlation citation string is stored in \srcref{vermicomposter_air_heater_design_model.m:L9-L10}. Equation numbering was not verifiable from repository contents alone.
\item [R3] Churchill friction-factor citation string is stored in \srcref{vermicomposter_air_heater_design_model.m:L11-L12}.
\item [R4] Churchill--Chu natural-convection citation string is stored in \srcref{vermicomposter_air_heater_design_model.m:L13-L15}.
\item [R5] The repository cites Bergman, Lavine, Incropera, and DeWitt generically for thermal resistances, developing internal flow, lumped balances, and free convection \srcref{vermicomposter_air_heater_design_model.m:L16-L17}. Edition and page numbers are not stored in the repository.
\item [R7] The helix geometry relation is documented as a repository-native design relation in comments and TeX narrative \srcref{vermicomposter_air_heater_design_model.m:L1708-L1727}; \srcref{Heat Transfer Study/VermiHeatLoss.tex:L38-L121}.
\item [R9] IAPWS saturation ancillary relations are cited in the repository commentary for water saturation properties \srcref{vermicomposter_air_heater_design_model.m:L18-L36}; \srcref{Heat Transfer Study/VermiHeatLoss.tex:L1089-L1167}.
\item [R10] The Lewis relation is cited in the repository commentary for mass-transfer closure \srcref{Heat Transfer Study/VermiHeatLoss.tex:L1174-L1180}. No separate source PDF was found in the repository tree.
\item [R11] Frederickson et al.\ Table~3 particle-size fractions are reproduced in the repository narrative \srcref{Heat Transfer Study/VermiHeatLoss.tex:L1015-L1088}. The repository does not include the original paper PDF.
\item NOAA and EIA URLs used by the annual layer are preserved in the configuration and narrative, not in a dedicated bibliography database \srcref{Heat Transfer Study/VermiHeatLoss.tex:L2574-L3164}; \srcref{Heat Transfer Study/study_config.json:L289-L419}.
\end{itemize}

\section*{Implementation topic map}
\begingroup
\small
\renewcommand{\arraystretch}{1.08}
\begin{longtable}{@{}L{0.16\textwidth}L{0.28\textwidth}L{0.48\textwidth}@{}}
\toprule
Concept & Governing equations and source status & Implementation mapping and notes \\
\midrule
\endfirsthead
\toprule
Concept & Governing equations and source status & Implementation mapping and notes \\
\midrule
\endhead
Tube flow, perforations, electrical grouping &
Equations: \code{eq:volflow\_per\_tube}--\code{eq:electrical\_totals}. \newline
Source status: repository derivation. &
Implementation: \srcref{vermicomposter_air_heater_design_model.m:L763-L875}; \srcref{vermicomposter_air_heater_design_model.m:L1497-L1677}. \newline
Notes: series-parallel grouping allowed. \\
Internal convection and friction &
Equations: \code{eq:internal\_laminar\_nu}, \code{eq:internal\_turbulent\_nu}, \code{eq:churchill\_friction\_full}. \newline
Source status: [R2], [R3], [R5]. &
Implementation: \srcref{vermicomposter_air_heater_design_model.m:L2382-L2407}; \srcref{heat_transfer_study.py:L3755-L3774}. \newline
Notes: used in both heater and aeration branches. \\
Heater axial energy balance &
Equations: \code{eq:heater\_segment\_air}, \code{eq:wire\_delta\_t}. \newline
Source status: [R1], [R5] plus repository discretization. &
Implementation: \srcref{vermicomposter_air_heater_design_model.m:L1803-L1894}; \srcref{heat_transfer_study.py:L3775-L3895}. \newline
Notes: explicit marching. \\
Aeration wall/jet/latent coupling &
Equations: \code{eq:aeration\_q\_wall}--\code{eq:aeration\_air\_march}. \newline
Source status: repository reduction of sensible plus latent paths. &
Implementation: \srcref{vermicomposter_air_heater_design_model.m:L1895-L1999}; \srcref{heat_transfer_study.py:L2815-L2959}. \newline
Notes: same structure in heating and cooling. \\
Evaporation and latent heat &
Equations: \code{eq:sauter\_diameter}--\code{eq:evap\_selected}. \newline
Source status: [R1], [R9], [R10], [R11]. &
Implementation: \srcref{vermicomposter_air_heater_design_model.m:L891-L937}; \srcref{vermicomposter_air_heater_design_model.m:L2000-L2122}; \srcref{heat_transfer_study.py:L2567-L2603}; \srcref{heat_transfer_study.py:L3607-L3708}. \newline
Notes: minimum of three limiting rates. \\
Bin loss model and two-node bed &
Equations: \code{eq:topmode\_open}--\code{eq:two\_node\_split}. \newline
Source status: [R4], [R5] plus repository vented-cover extension. &
Implementation: \srcref{vermicomposter_air_heater_design_model.m:L971-L1171}; \srcref{heat_transfer_study.py:L4326-L4438}. \newline
Notes: lumped bed closure. \\
Cooling spot cooler and assist blower &
Equations: cooling and condensate relations in Sections 12--15. \newline
Source status: repository surrogate model. &
Implementation: \srcref{heat_transfer_study.py:L2534-L3292}. \newline
Notes: includes condensate-limited RH update. \\
Heating and cooling NSGA-II &
Equations: \code{eq:heating\_constraint\_residuals}--\code{eq:cooling\_report\_selection}. \newline
Source status: repository implementation. &
Implementation: \srcref{heat_transfer_study.py:L51-L121}; \srcref{heat_transfer_study.py:L2181-L2499}; \srcref{heat_transfer_study.py:L3148-L3208}; \srcref{heat_transfer_study.py:L6220-L6490}. \newline
Notes: cooling now also NSGA-II. \\
Annual climate and duty layer &
Equations: \code{eq:kernel\_baseline}, \code{eq:daily\_duty}, \code{eq:daily\_energy}. \newline
Source status: repository numerical construct plus NOAA/EIA inputs. &
Implementation: \srcref{heat_transfer_study.py:L4633-L5012}. \newline
Notes: not a literature first-principles climate model. \\
\bottomrule
\end{longtable}
\endgroup

\end{document}
